---
title: "Midterm 3"
author: "Joyce Gill, Khanh Do, Matthew Billings-Chiu"
date: "2025-04-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(DBI)
library(odbc)
library(RMySQL)
library(tidyverse)
library(dbplyr)
library(RSQLite)
```

swipe time, swipe ratio, app usage time label APP USAGE METRICS

gender, sexual orientation, liked received DOES GENDER/SEX RELATE TO LIKES RECEIVED

likes receivd, swipe right label, mutual matches and match outcome DOES NUM LIKES and HOW OFTEN U SWIPE GIVE GOOD MATCH RESULT

## Read Data
```{r}
## Data
data <- read.csv("https://raw.githubusercontent.com/khanhdo05/stats-midterm-3/refs/heads/main/dating_app_behavior_dataset.csv")
data <- data %>% mutate(ID = row_number())
head(data)
```

## Connect to Database
```{r}
mydb <- dbConnect(RSQLite::SQLite(), ":memory:")
```

## Khanh 

```{r}
# gender, sexual orientation, likes received DOES GENDER/SEX RELATE TO LIKES RECEIVED
# dbExecute(mydb, "DROP TABLE `dating`")
dbExecute(mydb, "CREATE TABLE `gender_related_to_sex` (
  `ID` int(11) NOT NULL,
  `gender` varchar(255) NOT NULL,
  `sexual_orientation` varchar(255) NOT NULL,
  `likes_received` int(11) NOT NULL
)")

gender_related_to_sex_data <- data %>%
  select(ID, gender, sexual_orientation, likes_received)

dbWriteTable(conn = mydb, name = 'gender_related_to_sex', value = gender_related_to_sex_data, append = TRUE, header = FALSE, row.names = FALSE)

## Check
dbGetQuery(mydb,"Select * from gender_related_to_sex LIMIT 5;")
```

## Matthew

likes received, swipe right label, mutual matches and match outcome
```{r}
dbExecute(mydb, "CREATE TABLE `net_outcome` (
  `likes_received` int(11) NOT NULL,
  `swipe_right_label` varchar(15) NOT NULL,
  `mutual_matches` int(2) NOT NULL,
  `match_outcome` varchar(30) NOT NULL,
  `ID` int(5) NOT NULL
)")

net_outcome_data <- data %>%
  select(likes_received, swipe_right_label, mutual_matches, match_outcome, ID)
```

```{r}
## Fill Table
dbWriteTable(conn = mydb, name = 'net_outcome', value = net_outcome_data, append = TRUE, header = FALSE, row.names = FALSE)

## Check
dbGetQuery(mydb,"Select * from net_outcome LIMIT 5;")
```



```{r}
dbExecute(mydb, "CREATE TABLE `dating` (
  `app_usage_time_min` int(5) NOT NULL
)")
```

```{r}
dbWriteTable(conn = mydb, name = 'dating', value = data, append = TRUE, header = FALSE, row.names = FALSE)

## Check
dbGetQuery(mydb,"Select * from dating LIMIT 5;")
```



