---
title: "Midterm 3 Report: Managing a SQL Database"
author: "Joyce Gill, Khanh Do, Matthew Billings-Chiu"
date: "2025-04-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(DBI)
library(odbc)
library(RMySQL)
library(tidyverse)
library(dbplyr)
library(RSQLite)
```

# Data of Interest

Our group chose a dating app behavior dataset because of its relevance in today’s tech-driven social interactions. It offers fascinating insights into user behavior, preferences, and match outcomes across gender, activity levels, and time-of-day usage patterns. This can inform better app design and user experience, which aligns with interests in data science, behavior analysis, and software product improvement.

```{r}
## Data
data <- read.csv("https://raw.githubusercontent.com/khanhdo05/stats-midterm-3/refs/heads/main/dating_app_behavior_dataset.csv")
```

```{r}
## Connect to Database
mydb <- dbConnect(RSQLite::SQLite(), ":memory:")
```

```{r}
# Subtable: Gender-Sexual Orientation Pair
gender_so_table <- data %>%
  distinct(gender, sexual_orientation) %>%
  arrange(gender, sexual_orientation) %>%
  mutate(gender_sexual_orientation_id = paste(gender, sexual_orientation, sep = "-"))

# View the table
print(gender_so_table)

# Subtable: Location Type-Income Bracket Pair
location_income_table <- data %>%
  distinct(location_type, income_bracket) %>%
  arrange(location_type, income_bracket) %>%
  mutate(location_type_income_bracket_id = paste(location_type, income_bracket, sep = "-"))

# View the table
print(location_income_table)

# Make subtable in SQL
dbExecute(mydb, "CREATE TABLE `gender_so_label` (
  `gender_sexual_orientation_id` varchar(255) NOT NULL,
  `gender` varchar(255) NOT NULL,
  `sexual_orientation` varchar(255) NOT NULL
);")

dbWriteTable(conn = mydb, name = 'gender_so_label', value = gender_so_table, append = TRUE, header = FALSE, row.names = FALSE)

dbGetQuery(mydb, "SELECT * FROM gender_so_label")

dbExecute(mydb, "CREATE TABLE `location_income_label` (
  `location_type_income_bracket_id` varchar(255) NOT NULL,
  `location_type` varchar(255) NOT NULL,
  `income_bracket` varchar(255) NOT NULL
);")

dbWriteTable(conn = mydb, name = 'location_income_label', value = location_income_table, append = TRUE, header = FALSE, row.names = FALSE)

dbGetQuery(mydb, "SELECT * FROM location_income_label")
```
```{r}
## Table 1: Demographics
data <- data %>%
  left_join(
    gender_so_table %>%
      dplyr::select(gender, sexual_orientation, gender_sexual_orientation_id),
    by = c("gender", "sexual_orientation")
  ) %>%
  left_join(
    location_income_table %>%
      dplyr::select(location_type, income_bracket, location_type_income_bracket_id),
    by = c("location_type", "income_bracket")
  ) %>%
  distinct()

demographics_data <- data %>%
  dplyr::select(
    gender_sexual_orientation_id,
    location_type_income_bracket_id,
    education_level,
    interest_tags
  )

dbExecute(mydb, "CREATE TABLE `demographics` (
  `gender_sexual_orientation_id` varchar(255) NOT NULL,
  `location_type_income_bracket_id` varchar(255) NOT NULL,
  `education_level` varchar(255) NOT NULL,
  `interest_tags` varchar(255) NOT NULL
);")

dbWriteTable(conn = mydb, name = "demographics", value = demographics_data, append = TRUE, row.names = FALSE)

dbGetQuery(mydb, "SELECT * FROM demographics LIMIT 5;")
```

```{r}
## Subtable: app_usage_time_label
usage_range_table <- data %>%
  group_by(app_usage_time_label) %>%
  summarise(
    min_time = min(app_usage_time_min, na.rm = TRUE),
    max_time = max(app_usage_time_min, na.rm = TRUE)
  ) %>%
  arrange(min_time) %>%
  mutate(
    range = paste0(min_time, "-", max_time)
  ) %>%
  dplyr::select(range = range, label = app_usage_time_label)

# View the table
print(usage_range_table)  

## Subtable: swipe_right_label
swipe_right_range_table <- data %>%
  group_by(swipe_right_label) %>%
  summarise(
    min_time = min(swipe_right_ratio, na.rm = TRUE),
    max_time = max(swipe_right_ratio, na.rm = TRUE)
  ) %>%
  arrange(min_time) %>%
  mutate(
    range = paste0(min_time, "-", max_time)
  ) %>%
  dplyr::select(range = range, label = swipe_right_label)

# View the table
print(swipe_right_range_table)  

# Make subtable in SQL
dbExecute(mydb, "CREATE TABLE `app_usage_time_range_label` (
  `range` varchar(255) NOT NULL,
  `label` varchar(255) NOT NULL
);")

dbWriteTable(conn = mydb, name = 'app_usage_time_range_label', value = usage_range_table, append = TRUE, header = FALSE, row.names = FALSE)

dbGetQuery(mydb,"Select * from app_usage_time_range_label")

dbExecute(mydb, "CREATE TABLE `swipe_right_range_label` (
  `range` varchar(255) NOT NULL,
  `label` varchar(255) NOT NULL
);")

dbWriteTable(conn = mydb, name = 'swipe_right_range_label', value = swipe_right_range_table, append = TRUE, header = FALSE, row.names = FALSE)

dbGetQuery(mydb,"Select * from swipe_right_range_label")
```
```{r}
## Table 2: App Usage Metrics
```

```{r}
## Table 3: Dating app techniques: 
## likes_received, mutual_matches, profile_pics_count, bio_length, 
## message_sent_count, emoji_usage_rate, last_active_hour, swipe_time_of_day, match_outcome
dbExecute(mydb, "CREATE TABLE `dating_app_techniques` (
  `likes_received` int(25) NOT NULL,
  `mutual_matches` int(25) NOT NULL,
  `profile_pics_count` int(25) NOT NULL,
  `bio_length` int(25) NOT NULL,
  `message_sent_count` int(25) NOT NULL,
  `emoji_usage_rate` dbl(25) NOT NULL,
  `last_active_hour` int(25) NOT NULL,
  `swipe_time_of_day` varchar(255) NOT NULL,
  `match_outcome` varchar(255) NOT NULL
);")

dating_app_techniques_data <- data %>%
  dplyr::select(
    likes_received, mutual_matches, profile_pics_count, bio_length, 
    message_sent_count, emoji_usage_rate, last_active_hour, 
    swipe_time_of_day, match_outcome
    )

dbWriteTable(conn = mydb, name = 'dating_app_techniques', value = dating_app_techniques_data, append = TRUE, header = FALSE, row.names = FALSE)

dbGetQuery(mydb,"Select * from dating_app_techniques LIMIT 5;")
```

2. app usage metrics number: app usage time min, swipe right ratio
  subtable for app usage time label, app usage time min (turned into a range), 
  subtable for swipe right label, swipe right ratio (turned into a range)
  

```{r}
dbDisconnect(mydb)
```